---
title: "Assignment 3 - Looping Car Power Analyses"
author: "Evie Vermeer"
date: "3/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#attach packages
library(tidyverse)
library(purrr)
library(ggpubr)
library(ggplot2)
```

* Add an additional super light car with mass 5,000 kg and a surface area 10m2

* Compute the mean and maximum power for Cars A, B, C and for your new car, assuming that average highway speed is 80 km/hr with a standard deviation of 10km/hr;  Use 2 different methods to compute the mean and maximum power for each car type; a) **FOR** loop and b) the **pmap** function from **purr**

* Create two boxplots (one for each method (**FOR** and **pmap**)) that show the range of power consumption (across sampled speeds) for each car type.

```{r}
#source function for auto power 
source("~/Bren/Q2/ESM 262 Computing for Environmental Science/examples/ESM_262_Examples/R/autopower.R")

```

For this analysis we will use four different car types, a range of speed data and the auto power function to calculate the power required to maintain various speeds 

* car A  mass 20000 kg, area 22 m^2^

* car B mass 65000 kg, area 30 m^2^

* car C mass 38000 kg area 22 m^2^

* car D mass 5000 kg and area 10m^2

Mean highway speed of data is 80 km/hr with standard deviation of 10 km/hr 


```{r}
#create dataframe for cars 

possible_cars = data.frame(name = c("A","B","C","D"),mass=c(10000,65000,38000,5000), area = c(22,30,22,10))

```

```{r}
#create speed data sample
nsample = 200
mean_speed = 80

speeds = rnorm(mean=mean_speed, sd=10, nsample)
summary(speeds)


#plot speed data to verify output
plot(density(speeds), ylab="Distribution of Speeds in (m/s)")
```
```{r}
#run pmap loop - various loop data are car areas and masses, using the autopower function and different speeds
powercarspmap = pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds)

#add values to dataframe. Name columns after the car names 
powercarspmap = as.data.frame(powercarspmap, col.names=possible_cars$name)

#turn columns to rows for easier plotting later on 
powercarspmap = powercarspmap %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
```



```{r}
# Before running the for loop, create a data frame that has the # of rows = the # of speed data. The # of columns = the # of cars
powercarsfor = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))

#add column names 
colnames(powercarsfor)=possible_cars$name

#run the for loop and add values to dataframe
for (i in 1:nrow(possible_cars)) {
  powercarsfor[,i] = autopower(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}

#turn columns to rows for easier plotting 
powercarsfor = powercarsfor %>% pivot_longer(cols=everything(), names_to="car", values_to="power")

```


```{r}
#plotting - create two different plots based on your data 

pmap_plot = ggplot(powercarspmap, aes(car,power, fill=car))+geom_boxplot()+ggtitle("PMap Loop")

for_plot = ggplot(powercarsfor, aes(car,power, fill=car))+geom_boxplot()+ggtitle("For Loop")

#use the ggarrange tool to overlay 
ggarrange(pmap_plot, for_plot, nrow=2)

```

